Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-02-11 bbbq_article.tex
Use Bogaart-Shakespearean, fix #177
Discuss haplotype that presents TMH-derived epitopes from SARS-CoV-2 four times as much, thanks @fransbianchi, fix #71

We found that the percentages of epitopes overlapping 
with TMHs for a human and viral proteome are 
[similar/different]. In other words, the
epitopes that MHC-I presents are [as/not as] likely 
to be derived from TMH within either a human host and its viral pathogen.

We found that the percentages of epitopes overlapping 
with TMHs for a human and bacterial proteome are 
[similar/different]. In other words, the
epitopes that MHC-I presents are [as/not as] likely 
to be derived from TMH within either a human host and its bacterial pathogen.

@@

We concluded that the
epitopes that MHC-I presents are [as/not as] likely 
to be derived from TMH within either a human host and its viral pathogen.
Because the full SARS-CoV-2 has 21 TMHs, the percentages
of MHC-I epitopes being part of a TMH are likelier to be affected by
stochasticity. We chose to use SARS-CoV-2 regardless, as the thousands
of its time-dated genomic sequences are ideal for determining the 
evolutionary conservation of MHC-I detecting TMHs.

@@

We used a low number of SNPs, as was imposed by the
NCBI API, which constrains its users to three calls per second.
As a second factor, we limited the time querying the NCBI database
for the duration of the project.
The SNPs we did use, we picked those
with the lowest SNP ID (which denotes the earliest discovery date) first.
We expect this not to introduce a bias, so we expect that
the SNPs we used are a representative sample of all SNPs.

\paragraph{Elution studies}

\paragraph{\% epitopes overlapping with TMHs}

We have removed the 25 selenoproteins found in the human
reference proteome, as both prediction tools struggle with 
selenocysteine, the 21st amino acid incorporated in human proteins.
TMHMM gives a clear error message, forcing us to remove
these selenoproteins from the analysis.
How PureseqTM deals with selenocysteine is unclear,
as it gives no error, nor is it mentioned anywhere,
nor do the authors respond to emails. Due to this, we removed the
selenoproteins from our analysis.
Another bug in PureseqTM is that it does not predict the topology
of proteins that have less than three amino acids. 
The TRDD1\_HUMAN protein, short for 'T cell receptor delta diversity 1',
is, however, two peptides long. We manually added the prediction
that both peptides reside in the cytosol.

\paragraph{Evolutionary conservation}

Translational staling.

Trans-splicing for epitopes.

\richel{
  Note that most bacteria are opportunistic pathogens.
  Note that most bacteria are generalists.
  Note that most bacteria have different cell membranes (and walls), that
  may have different functional constraints than a human cell membrane
}

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_snps_found_and_expected_per_n_tmhs.png}
  \caption{
    Number of SNPs found in TMHs,
    separated per number of TMHs.
    The dashed diagonal lines show the expected linear trend line
    if TMHs and soluble protein regions are equally conserved.
  }
  \label{fig:n_snps_found_and_expected_per_n_tmhs}
\end{figure}

@@

>>>>

\paragraph{General pattern in epitope presentation}

We found a general pattern in the
presentation of epitopes that overlap with a TMH.
For humans, a bacterial and viral pathogen,
as well as for MHC-I and MHC-II,
we found that these are presented more often 
than expected by chance.

\paragraph{Interesting haplotype for SARS-CoV-2}

One interesting finding is that the MHC-I 
haplotype HLA-A*24:02 
present four times as much epitopes derived from SARS-CoV-2 TMHs
than expected by chance.
This may help the immune system to detect a SARS-CoV-2 infection earlier,
but we have not investigates if an effect can be found in clinical data.

@@

\subsection{Elution studies}

\paragraph{False positives}

\subsection{\% epitopes overlapping with TMHs}

\paragraph{Removal of selenoproteins}

We have removed the 25 selenoproteins found in the human
reference proteome, as both prediction tools struggle with 
selenocysteine, the 21st amino acid incorporated in human proteins.
TMHMM gives a clear error message, forcing us to remove
these selenoproteins from the analysis.
How PureseqTM deals with selenocysteine is unclear,
as it gives no error, nor is it mentioned anywhere,
nor do the authors respond to emails. Due to this, we removed the
selenoproteins from our analysis.

\paragraph{Topology prediction of TRDD1}

PureseqTM does not predict the topology
of proteins that have less than three amino acids. 
The TRDD1 ('T cell receptor delta diversity 1') protein,
however, is two peptides long. 
We corrected for this by predicting that this protein
is completely cytosolic.

\paragraph{Bacteria have a different cell membrane}

In this experiment we predicted epitopes that overlap with 
TMHs from a human, bacterial and viral proteome,
would these proteins be expressed in a human host.
Bacteria, however have different cell membranes and cell walls, 
hence different structural requirements for a TMH.
Both topology prediction tools were trained to recognize
human TMHs (\richel{Check}), thus we cannot be sure that
the transmembrane regions predicted in bacterial proteins
are actually part of an integral membrane protein.
For the purpose of this study, we assume the 
error in topology predictions to be unbiased way towards topology.
In other words: that a bacterial TMH is incorrectly
predicted to be absent just as often as it is incorrectly
predicted to be present elsewhere.

\subsection{Evolutionary conservation}

\paragraph{synonymous mutations}

\paragraph{False positives in SNPs}

\paragraph{Low number of SNPs}

The NCBI dbSNP database contains millions of SNPs.
As the NCBI API constrains its users to three calls per second,
we had to limited the extent of our analysis
by downloading SNPs to approximately 300 
hours (i.e. the length of a Christmas holiday).

\paragraph{Selection of SNPs}

We selected to analyze SNPs with the lowest SNP ID first.
This means we have a bias for picking SNPs with
an earlier discovery date.
We expect this bias to be biologically irrelevant
and we find no clear evidence of a bias, as
shown in figure \ref{fig:n_tmhs_per_protein}.
Still, the best way to interpret this figure
is open to discussion.

\subsection{Notes and/or to be removed}

@@

@@

@@

% Instead of non-TMH use 'soluble protein region'
% or 'cytosolic or extracellular protein region'
